Prey movement shapes the development of predator expertise in a virtual bi-trophic system

### 2024-06-05 manuscript/manuscript.Rmd
Ms up to date. discussion à finir

The ecological consequences of predation are defined by the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [refs]. Recent advances in our understanding of the behavioural mechanisms (e.g. reciprocal plasticity, foraging specialization) driving prey consumption at the individual and population level have allowed ecologists to rely on foraging strategies to predict the outcome of (N)CEs and their effect on prey population dynamics [refs]. For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007]. Because ambush predators are opportunistic hunters, they also tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].

While the foraging mode of predators is a reliable predictor of prey consumption, predators need to acquire expertise by practicing and learning the proper skills to successfuly capture their prey [@dukas2019;wooster.etal2023]. The aquisition of expertise is predicted to be paramount in many predator systems where hunting tactics are learned (e.g. orca whales). Yet, there is only fraction of evidence showing the link between experience and success in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.

Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) by learning associative images or using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015]. Predators can integrate this information to use the appropriate hunting tactic (e.g. ambushing or fast movement speeds) for the type of prey that they encounter [@toscano2014;@patrick.weimerskirch2014a;@toscano.etal2016]. However, prey often use antipredator tactics such as rapid or unpredictable movements which are effective to successfully escape predators [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022]. This results in reciprocal interactions where predators need to learn to predict the movement of their prey to capture them. Therefore, antipredator tactics may impede the acquisition of hunting expertise by interfering with learning processes and reducing the success of predators.

A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey. Similar to agent-based simulations, online videogames can be useful systems to mitigate these challenges and test specific ecological hypotheses. They provide controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023] with the advantage of having real players that interact in the virtual space. For example, in the predator-prey videogame *Dead by Daylight*, four prey players need to forage for resources while avoiding predation by a fifth player. In this virtual system, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e. mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022]. The prey can increase their chances of survival by cooperating and moving fast when they forage [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024]. The game also ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020]. These observations outline how ecological patterns can emerge from interactions in virtual systems with fixed rules. Videogames also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios. Hence, virtual systems such as *Dead by Daylight* offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.

In this study, we assess how repeated experience shapes predator foraging success using data from players in *Dead by Daylight*. First, we investigate how individual hunting success changes with experience (i.e. how individuals develop their hunting expertise). We hypothesize that the predators' success will increase with experience up to a certain level where it will stagnate. We expect this pattern to vary among individuals due to differences in prey encounters. Second, we investigate how prey movement influences the development of expertise. If all individual predators encounter groups of prey with similar speeds, we predict that their gain in expertise will be similar.

In contrast, based on previous results showing that faster prey are more difficult to hunt, then...

if all predators encounter varying groups of prey, then they should all converge towards flexible speeds. In both scenarios, differences among individuals in IIV across experience should be low (i.e. similar individual foraging specialization), whereas the population variance would either decrease (specialization) or increase (flexibility).

@@

*DBD* is a survival asymmetric multiplayer online game (i.e. a game where the gameplay mechanics differ between two groups) developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey. Every match includes only one predator and four prey. The objective of the predator is to hunt and capture the prey, and the objective of the prey is to search for resources while avoiding the predator. The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors. The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm. A match ends when the predator kills all the prey available (i.e. that have not escaped), or when the last remaining prey escapes the virtual environment. Each player, predator or prey, can choose an avatar with abilities that encourage specific play styles (e.g. bold vs cautious prey, or ambush vs roaming predator). During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars. The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings. Some of these environments are larger than others, with varying structural complexity. However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues of the generators [@fraserfranco.etal2022]. There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.

@@

Our population consists of `r toString(format(length(unique(dat$predator_id))))` predator players with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches. The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played. These matches lasted between `r toString(format(round(min(dat$game_duration / 60))), digits = 2)` and `r toString(format(round(max(dat$game_duration / 60))), digits = 2)` min (mean = `r toString(format(round(mean(dat$game_duration / 60))), digits = 2)` min). The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e. the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey the player encountered, and their mean rank (a proxy for prey skill). The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.

We analyzed the mean speed of the prey group encountered by the predator. We measured the preys' speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s). We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`). Lastly, we defined the predator's cumulative experience as the number of matches played as the predator prior to the match being monitored. For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9. We did not account for matches where predators played as the prey. We also categorized predators for each match based on their cumulated experience. We labeled predators as novices for matches where they had cumulated less than 100 matches, intermediate for matches where they had cumulated between 100 and 299 matches, and advanced for matches where they had cumulated more than 299 matches (max 499). Since our goal was to monitor predator players throughout their experience and that they all played at least 300 matches, they all appeared in each of the three experience categories.

We recognize that we could have biased our analyses by retaining only individuals who played for at least 300 matches. For example, these individuals might be experienced videogame players and could thus already be playing like experts in their first matches in *DBD*. To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population. We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean speed using a Bayesian hierarchical linear model. We found that all four groups had similar mean speeds as predators (Appendix 2: Table S1-S2), which indicates an absence of bias due to data sampling.

@@

@@

All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). Each required 64GB of RAM with 48 cores to compile within 5 days.

We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parametrized the models following the method of @pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were reproductions of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience. We included the standardized match duration and prey rank as covariates in all five models.

We computed the five models using a modified version of the beta-binomial distribution implemented in "brms". Hunting success was estimated as the probability of consuming the four prey ($\mu_i$), drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters. We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor, while the precision parameter ($\phi$) was estimated with an identity link. We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience. We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$). We used weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$). Following @fraserfranco.etal2022, we defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$), a positive Gaussian prior ($N(1,\ 0.5^{2}))$) on the game duration because longer trials lead to greater success, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system. We employed weakly informative half-Gaussian priors on all the standard deviation parameters ($N(0,\ 0.5^{2})$). We compared the models' predictive accuracies using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].

We parametrized the GAMMs and the MDHGLM to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models. We set the first 500 iterations in each model as warm ups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics). We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.

@@

Out of all five GAMM models, two predicted the data the best and achieved similar expected log pointwise densities (elpd). Both models accounted for the prey group's rank and speed (Table 1). Models in which these effects were not accounted for resulted in no change in hunting success with experience (i.e. expertise) for the average individual (results not shown). Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A). In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve, with predators optimizing their success after playing ~300 matches (Figure 1C). The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).

@@

The relationship between hunting success and cumulative experience differed among predators (Figure 1B-D). Only `r toString(scales::percent(percent_tab1[2,]$percentage, accuracy = 0.1))` of the population had an increase in success from the first match to the last in the model where we did not account for the prey's speed and rank (Appendix S1: Figure S1A). In contrast, accounting for both effects resulted in `r toString(scales::percent(percent_tab1[5,]$percentage, accuracy = 0.1))` of the population increasing its success with experience from the first match to the last. The prey speed alone also mediated individual differences in the relationship between success and experience. The standard deviation of the individual slopes component of the model accounting only for prey rank was equal to 9.72 (9.30, 10.15 $95\%\ CI$), while the one for the model accounting for prey rank and speed was equal to 3.36 (3.04, 3.71 $95\%\ CI$), indicating that individual differences in the development of expertise decreased when accounting for prey speed.

@@

A general assumption of predator-prey studies is that predators maximize success by matching their tactic to their prey [@abrams2000], yet, it has remained largely unknown whether this results from predators learning how to hunt their prey in part because of the challenges of investigating direct interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that predators in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice. **parler de l'effet des proies**

## The development of expertise

Our results suggest that predator expertise is honed through extensive practice. The predator population increased and stabilized its success with experience in a diminishing returns fashion as it is typically found in empirical studies of expertise [reviewed in @dukas2019]. The prey were important in mediating this pattern, probably because they increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This resulted in discernable differences in the relationship between success and experience among predators, implying that the development of expertise depends on an individual's capacity to adjust to difficulty. Hunting faster prey requires costly and specialized cognitive abilities and coordination [@kelley.magurran2011], and that could explain why such prey were more difficult to capture. Thus, predators that couldn't properly hunt at high speeds or develop counter-strategies for faster prey were likely at a disadvantage.

## The role of prey behaviour

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predators, suggesting that other antipredator tactics were potentially involved. For example, if the prey used camouflage or hiding to avoid being detected, then predators may have found them only if they learned how to exploit the visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020].

In addition, longer time intervals between hunting events may delay or even impede learning due to forgetting important information [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.

Yet, because prey can also learn how to avoid predation [@kelley2003;@turner2006;@mccomb2011], we believe that the predator-prey phenotype matching more likely emerged from reciprocal adjustments in speed by predators and prey as they interacted [@kishida.etal2006;@kishida.etal2009;@edgell.rochette2009;@mcghee.etal2013]. Hence, if the prey also learned through repeated interactions with the predators, it is possible that experience contributed in stabilizing the system as both were adjusting to each other, similar to Red Queen dynamics [@brockhurst2014].

However, there were still considerable differences in success among individuals through time, suggesting that some predators were limited in their capacity to match their tactic to their prey or to increase their success through other means. Specialist foragers were faster, and thus, probably better equipped to hunt the more difficult faster prey in *DBD*. However, if the prey responded to fast predators by also being faster, then hunting at high speeds resulted in more difficult encounters for these predators, thereby decreasing the benefits of using this tactic (Figure 4). Thus, specializing probably compensated for the difficulty of hunting prey at high speeds by helping predators to better predict the location and movement of their prey. On the other hand, flexible foragers encountered variable prey with slower speeds. Yet, because the prey increased their speed with experience, the benefits of being able to hunt multiple prey types for flexible hunters may have come at the cost of not being adept at capturing faster prey [@pintor.etal2014]. Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.

@@

>>>>

The ecological consequences of predation accrue to the sum of consumptive and nonconsumptive effects (CEs and NCEs) during predator-prey interactions [@preisser2007]. Recent advances in our understanding of the behavioural mechanisms (e.g. foraging mode, foraging specialization, reciprocal plasticity) driving (N)CEs at the individual and population level have allowed ecologists to predict their outcome as well as their effect on prey population dynamics [@wirsing.etal2021]. For example, empirical studies show that prey tend to increase vigilance at the expense of foraging in the presence of ambush predators (NCEs) [@preisser2007]. Because ambush predators are opportunistic hunters, they tend to consume a greater range of prey types (CEs) compared to cursorial predators [@glaudas2019].

There are, however, internal mechanisms shaping predator foraging decision-making that are often overlooked when studying (N)CEs in predator-prey interactions [@wooster.etal2023]. For instance, predators need to acquire expertise by practicing and learning the proper skills to successfully locate, select, and capture their prey [@dukas2019;@wooster.etal2023]. The aquisition of expertise is predicted to be paramount in systems where predators learn different hunting tactics (e.g. orca whales) [@heithaus2018]. Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) by learning associative images or using cues from their prey and their environment [@edwards.jackson1994;@morse2000;@macdonald2007;@reid.etal2010;@wilson-rankin2015]. Thus, predators can integrate this information to increase their chances of locating prey or to use the tactic most suited for the type of prey encountered [refs].

Conversely, prey can use antipredator tactics that are effective to evade predators, such as rapid or unpredictable movements [@walker.etal2005;@kelley.magurran2011;@herbert-read.etal2017;@szopa-comley.ioannou2022]. This results in reciprocal interactions where predators need to learn to predict the movement of their prey to successfully capture them. Yet, it may be impossible for a predator to learn unpredictable prey escape trajectories such that these antipredator tactics may interfere with a predator's acquisition of hunting expertise [troscianko.etal2018]. Unfortunately, there is only scarce evidence showing links between experience, prey antipredator behaviour, and prey consumption in natural predators, impeding our ability to accurately predict the consequences of predation on prey populations.

A recurring challenge impeding research on predator-prey behavioural interactions is the need to collect data simultaneously on both the predator and the prey. Similar to agent-based simulations, online videogames can be useful systems to mitigate these challenges and test specific ecological hypotheses. They provide controlled virtual environments to test ecological hypotheses [see @lymbery.etal2023] with the advantage of having real players that interact in the virtual space. For example, in the predator-prey videogame *Dead by Daylight*, four prey players need to forage for resources while avoiding predation by a fifth player. In this virtual system, the predator population is composed of individuals that either ambush or hunt at high speeds (i.e. mean movement speed along a slow-fast continuum), and their success is driven by the movement of the prey [@fraserfranco.etal2022]. The prey can increase their chances of survival by cooperating and moving fast to escape the predator [@cere.etal2021;@fraserfranco.etal2022;@santostefano.etal2024]. The game also ellicits natural reactions in players such as fear from predation (personal observations), which corroborates with another virtual ecological study showing that predation drives individual variation in risk perception [@beauchamp2020]. These observations outline how ecological patterns can emerge from interactions in virtual systems with fixed rules. Videogames also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios. Hence, virtual systems such as *Dead by Daylight* offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.

In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*. First, we investigate how individual hunting success changes with repeated encounters (i.e. how individuals develop their hunting expertise). At the population level, we hypothesize that the predators' success will increase with experience up to a certain level where it will stabilize [@dukas2019]. However, we expect this pattern to vary among individuals due to differences in the movement of prey encountered. Therefore, we then investigate how prey movement influences the development of expertise. If predators encounter groups of prey with similar speeds, we predict that the gain in expertise will be similar among individuals. Alternatively, if predators encounter groups of prey with varying speeds, then the acquisition of expertise will vary among individuals.

@@

*DBD* is a survival asymmetric multiplayer online game (i.e. a game where the gameplay mechanics differ between two groups) developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey. Every match includes only one predator and four prey. The objective of the predator is to hunt and capture the prey, and the objective of the prey is to search for resources while avoiding the predator. The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors. The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm. A match ends when the predator kills all the prey available (i.e. that have not escaped), or when the last remaining prey escapes the virtual environment.

Before the start of a match, players (predator or prey) can choose an avatar with unique abilities that encourage specific play styles (e.g. bold vs cautious prey, or ambush vs roaming predator). During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars. The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings. Some of these environments are larger than others, with varying structural complexity. However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues on the resources [@fraserfranco.etal2022]. There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period. Details on the basic characteristics of predator avatars are available at https://deadbydaylight.fandom.com/wiki/Killers. Details on the size and structure of the different virtual environments are available at https://dbdmaps.com/ and https://deadbydaylight.fandom.com/wiki/Realms.

@@

Our population consists of `r toString(format(length(unique(dat$predator_id))))` players that played strictly as the predator, with a total record of `r toString(formatC(length(unique(dat$match_id)), big.mark = " "))` matches. The predator-players' experience varied between `r toString(format(min(dat$total_xp_pred)))` and `r toString(format(max(dat$total_xp_pred)))` matches played. These matches lasted between `r toString(format(round(min(dat$game_duration / 60))), digits = 2)` and `r toString(format(round(max(dat$game_duration / 60))), digits = 2)` min (mean = `r toString(format(round(mean(dat$game_duration / 60))), digits = 2)` min). The following information is collected and reported for every match : the player's anonymous ID, its avatar (i.e. the predator character chosen with its specific skill-gameplay mechanics), the game environment, the predator-player's experience, the mean speed of the groups of prey that the predator player encountered, and the mean rank of the prey encountered (a proxy for prey skill). The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.

We analyzed the mean speed of the prey group encountered by the predator. We measured the preys' speed as the mean travel speed of the four individual prey in a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s). We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`). Lastly, we defined the predator's cumulative experience as the number of matches played as the predator prior to the match being monitored. For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9. We did not account for matches where predators played as the prey.

We recognize that we could have introduced a bias in our analyses by retaining only individuals who played for at least 300 matches. For example, these individuals might be experienced videogame players and could thus already be playing like experts in their first matches in *DBD*. To verify that our sample was not biased, we compared a random sample of players that played either 20 to 50 matches, 51 to 100 matches, or 101 to 300 matches during the same timeframe as our sampled population. We then took the first 20 matches played by these players, including those from our sampled population, and compared their mean hunting success using a Bayesian hierarchical linear model. We found that all four groups had similar successes as predators (Appendix 2: Table S1-S2), which suggests an absence of bias due to data sampling.

@@

We tested how predators developed their expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parametrized the models following the method of @pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were reproductions of the second and third models respectively, where we included the prey speed to assess its effect on the relationship between success and experience. We included the standardized match duration and prey rank as covariates in all five models.

We computed the five models using a modified version of the beta-binomial distribution implemented in "brms". Hunting success was estimated as the probability of consuming the four prey ($\mu_i$), drawn from a Beta distribution ($Beta(\mu_i, \phi)$) with mean ($\mu \in [0,1]$) and precision ($\phi > 0$) parameters. We used a logit link function to estimate $\mu_i$ where $\mu_i = \frac{exp(\eta_i)}{exp(\eta_i) + 1}$ and $\eta_i$ is the linear predictor, while the precision parameter ($\phi$) was estimated with an identity link. We used the default number of basis functions (K) in "brms" for the models to estimate the relationship between hunting success and experience. We assumed that the random intercepts for the predator ID ($id$) followed a Gaussian distribution with estimated standard deviation ($id \sim N(0,\ I\ \sigma_{id}^{2})$). We used weakly informative Gaussian priors for the intercept ($N(0,\ 0.5^{2})$) and the global trend of cumulative experience ($N(0,\ 2^{2}))$). Following @fraserfranco.etal2022, we defined a positive Gaussian prior on the precision parameter ($N(2,\ 0.5^{2})$), a positive Gaussian prior ($N(1,\ 0.5^{2}))$) on the game duration because longer trials lead to greater success, and a negative Gaussian prior on prey speed ($N(-1,\ 0.5^{2})$) because encountering faster prey is associated with lower success in this system. We employed weakly informative half-Gaussian priors on all the standard deviation parameters ($N(0,\ 0.5^{2})$). We compared the predictive accuracy of all five models using approximate leave-one-out cross-validation with Pareto-smoothed importance sampling [@vehtari.etal2017;@piironen.vehtari2017;@vehtari.etal2022].

We parametrized the GAMMs and the MDHGLM to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models. We set the first 500 iterations in each model as warm ups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@vehtari.etal2021]. We also performed posterior predictive checks which showed an adequate fit of the models. We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.

@@

All models were fitted in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@gabry.cesnovar2021] as the back-end for parameter estimation (cmdstan installation version 2.28.2). The models were run on Cedar (Operating system: CentOS Linux 7), a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). Each required 64GB of RAM with 48 cores to compile within 5 days. For details on how to reproduce our analyses, please consult the GitHub repository of this project (https://github.com/quantitative-ecologist/predator-expertise).

@@

Out of all five GAMM models, the two that accounted for the prey group's rank and speed were the best at predicting the data, achieving similar expected log pointwise densities (Table 1). Models in which prey effects were not accounted for resulted in no change in hunting success with experience (i.e. expertise) for the average individual (**Appendix XX**). Accounting for the prey rank resulted in a concave-shaped relationship, with the highest success ranging between 100 and 400 matches (Figure 1A). In the model where we additionally included the prey's speed, the effect of experience on hunting success for the average individual followed a diminishing returns curve as we predicted, with predators optimizing their success after playing ~300 matches (Figure 1C). The curve shows there was a 38% increase in the probability of consuming all prey for the average individual between the first and the ~330$^{th}$ match where success reached a plateau (Figure 1C).

@@

The relationship between hunting success and cumulative experience differed among predators (Figure 1B-D). Only `r toString(scales::percent(percent_tab1[2,]$percentage, accuracy = 0.1))` of the population had an increase in success from the first match to the last in the model where we did not account for the prey's speed and rank (Appendix S1: Figure S1A). In contrast, accounting for both effects resulted in `r toString(scales::percent(percent_tab1[5,]$percentage, accuracy = 0.1))` of the population increasing its success with experience from the first match to the last. The prey speed mediated individual differences in the relationship between success and experience. The standard deviation of the individual slopes component of the model accounting only for prey rank was equal to 9.72 (9.30, 10.15 $95\%\ CI$), while the one for the model accounting for prey rank and speed was equal to 3.36 (3.04, 3.71 $95\%\ CI$), indicating that individual differences in the development of expertise decreased when accounting for prey speed.

@@

It is generally assumed that predators increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023]. Yet, few studies have empirically tested this hypothesis in part because of the challenges of investigating direct predator-prey interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience, suggesting that predator expertise was honed through extensive practice. Yet, there were important differences among individuals in the development of expertise, which was in part due to differences among individuals in the movement of the prey encountered.

Our results suggest that predator expertise is honed through extensive practice. The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated, consistent empirical observations of expertise in both humans and nonhuman animals [reviewed in @dukas2019] The prey's speed was important in mediating this pattern at the population level, probably because they also increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were probably encountering prey with higher speeds. Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.

Prey speed also mediated a portion of the differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise. Individual predator players probably differed in their capacity to adjust to difficult prey. For example, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011]. Thus, predators that couldn't develop counter-strategies for faster prey were likely at a disadvantage.

**Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.**

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predator players, suggesting that other antipredator tactics were potentially involved. For example, the prey can use camouflage and hide to avoid detection (not currently measurable in *DBD*), requiring predators to learn how to exploit visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020]. Furthermore, it is hypothesized that longer time intervals between hunting events could hinder or delay the acquisition of expertise as individuals may forget information with longer delays [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.

@@

**Comment by reviewer:**

Some characteristics to be aware of: 
- Predators do not compete.
- Both predators and prey can fail in a round yet still carry that experience and knowledge on to the next round with no real penalty (think of the movie Groundhog Day).
- Predator and prey are surely playing a game with each other, but prey may also be playing a game among themselves, which may affect their game with the predator. 
- This particular video game appears to have a few unusual characteristics.  Prey do not appear to forage for resources. but rather they create a public good.  Furthermore, handling time for the predator is extremely complicated and lengthy.
- Finally, it may be rather complicated to discern all of the rules of how the different virtual environments work and what  all the aptitudes of all the different predators and prey are.
